---
title: "Academic Indicators"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(reshape2)
library(directlabels)
```

## About

### Reading the data

```{r}
indicators <- read.csv(file="../Data/AcademicIndicators.csv", header=TRUE, sep=",", stringsAsFactors=FALSE)
str(indicators)
```

### Subsetting and melting the data - Why and How

```{r}
# Get only the columns we need
subset <- subset(indicators, select = c(IES,PP2013,PP2014,PP2015,PP2016))
# Change labels on the columns
colnames(subset) <- c("IES",2013,2014,2015,2016)
meltedsubset <- melt(subset, id.vars="IES", value.name="Count", variable.name="Year")
meltedsubset
```

### Setting up some layers and lists for ggplot

https://stackoverflow.com/questions/18747091/reuse-ggplot-layers-in-multiple-plots

http://www.cookbook-r.com/Graphs/Shapes_and_line_types/

```{r}
lthick <- c(1,1,1,2.5,1,1,1,1)
lstyle <- c("dashed","dashed","dashed","solid","solid","solid","dashed","dashed")
mygglayers = list(
    geom_line(aes(size=IES,linetype=IES)),
    geom_point(size=4, shape=21, fill="white"),
    scale_size_manual(values = lthick),
    scale_linetype_manual(values = lstyle)
)
```




## How many papers were published?

### Preprocessing and reshaping the data

```{r}
# Get only the columns we need
subset <- subset(indicators, select = c(IES,PP2013,PP2014,PP2015,PP2016))
# Change labels on the columns
colnames(subset) <- c("IES",2013,2014,2015,2016)
# Annotating the IES names with the grade in the last evaluation 
subset$IES = paste(subset$IES," (",indicators$Grade.in.2013,")",sep = "")
meltedsubset <- melt(subset, id.vars="IES", value.name="Count", variable.name="Year")
```

## The Plot

```{r fig.width=8, fig.height=4}
ggplot(data=meltedsubset, aes(x=Year, y=Count, group = IES, colour = IES)) +
    ggtitle("Papers") +
    mygglayers
```

## Adjusting for number of lecturers

https://stackoverflow.com/questions/39563283/r-divide-some-columns-of-a-data-frame-and-keeping-the-others

```{r fig.width=8, fig.height=4}
subset[,-1] = subset[,-1] / indicators[,c('Students2016')]
#subset[,-1] = subset[,-1] / indicators[,c('Lecturers2016')]
subset
meltedsubset <- melt(subset, id.vars="IES", value.name="Count", variable.name="Year")
ggplot(data=meltedsubset, aes(x=Year, y=Count, group = IES, colour = IES)) +
    ggtitle("Papers (divided by lecturers)") +
    mygglayers
```




## How many books were published?

### Preprocessing and reshaping the data

```{r}
# Get only the columns we need
subset <- subset(indicators, select = c(IES,BP2013,BP2014,BP2015,BP2016))
# Change labels on the columns
colnames(subset) <- c("IES",2013,2014,2015,2016)
meltedsubset <- melt(subset, id.vars="IES", value.name="Count", variable.name="Year")
```

```{r fig.width=8, fig.height=4}
ggplot(data=meltedsubset, aes(x=Year, y=Count, group = IES, colour = IES)) +
    ggtitle("Books") +
    mygglayers
```

## How many papers in conferences?

### Preprocessing and reshaping the data

```{r}
# Get only the columns we need
subset <- subset(indicators, select = c(IES,CN2013,CN2014,CN2015,CN2016))
# Change labels on the columns
colnames(subset) <- c("IES",2013,2014,2015,2016)
meltedsubset <- melt(subset, id.vars="IES", value.name="Count", variable.name="Year")
```

```{r fig.width=8, fig.height=4}
ggplot(data=meltedsubset, aes(x=Year, y=Count, group = IES, colour = IES)) +
    ggtitle("Papers in Conferences") +
    mygglayers
```

## How many talks?

### Preprocessing and reshaping the data

```{r}
# Get only the columns we need
subset <- subset(indicators, select = c(IES,TK2013,TK2014,TK2015,TK2016))
# Change labels on the columns
colnames(subset) <- c("IES",2013,2014,2015,2016)
meltedsubset <- melt(subset, id.vars="IES", value.name="Count", variable.name="Year")
```

```{r fig.width=8, fig.height=4}
ggplot(data=meltedsubset, aes(x=Year, y=Count, group = IES, colour = IES)) +
    ggtitle("Talks") +
    mygglayers
```


## How many short courses?

### Preprocessing and reshaping the data

```{r}
# Get only the columns we need
subset <- subset(indicators, select = c(IES,SC2013,SC2014,SC2015,SC2016))
# Change labels on the columns
colnames(subset) <- c("IES",2013,2014,2015,2016)
meltedsubset <- melt(subset, id.vars="IES", value.name="Count", variable.name="Year")
```

```{r fig.width=8, fig.height=4}
ggplot(data=meltedsubset, aes(x=Year, y=Count, group = IES, colour = IES)) +
    ggtitle("Short Courses") +
    mygglayers
```

## How many events were organized?

### Preprocessing and reshaping the data

```{r}
# Get only the columns we need
subset <- subset(indicators, select = c(IES,EO2013,EO2014,EO2015,EO2016))
# Change labels on the columns
colnames(subset) <- c("IES",2013,2014,2015,2016)
meltedsubset <- melt(subset, id.vars="IES", value.name="Count", variable.name="Year")
```

```{r fig.width=8, fig.height=4}
ggplot(data=meltedsubset, aes(x=Year, y=Count, group = IES, colour = IES)) +
    ggtitle("Events Organized") +
    mygglayers
```

## References


 (from https://stackoverflow.com/questions/17150183/plot-multiple-lines-in-one-graph)
 https://stackoverflow.com/questions/27350243/ggplot-line-graph-with-different-line-styles-and-markers
https://stackoverflow.com/questions/17996410/ggplot-specific-thick-line

labels:; factor order http://rstudio-pubs-static.s3.amazonaws.com/7433_4537ea5073dc4162950abb715f513469.html
