---
title: "Operations with Vectors"
output:   
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Basic Operations with Vectors

This is the monthly average temperature in Baltimore (from http://www.rssweather.com/climate/Maryland/Baltimore/)

```{r}
tempHi <- c(41.2,44.8,53.9,64.5,73.9,82.7,87.2,85.1,78.2,67.0,56.3,46.0)
tempHi
```

We can label the vectors' 
elements!
```{r}
months <- c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")
names(tempHi) <- months
tempHi
```


Other way to do this:
```{r}
otherTemp <- c(Jan=41.2,Feb=44.8,Mar=53.9,Apr=64.5,May=73.9,Jun=82.7,
               Jul=87.2,Aug=85.1,Sep=78.2,Oct=67.0,Nov=56.3,Dec=46.0)
otherTemp
names(otherTemp)
```

Let's also get the lower temperatures:

```{r}
tempLo <- c(23.5,26.1,33.6,42.0,51.8,60.8,65.8,63.9,56.6,43.7,34.7,27.3)
```

Now we can do simple operations on the whole vectors.

```{r}
range <- tempHi-tempLo
range
```

Convert it to Celsius:

```{r}
tempHiC <- (5/9) * (tempHi - 32)
tempHiC
```

Is this month too cold?

```{r}
tooCold <- (((5/9) * (tempLo - 32)) < 0)
tooCold
```

### Accessing Elements in a Vector

By indexes:

```{r}
tempHi[7] # July
tempHi[12] # December

```

By ranges:
```{r}
tempHi[7:9] # July-September
```

By lists:

```{r}
tempHi[c(11:12,1:2)] # Nov-Feb
```

By sequences:

```{r}
tempHi[seq(2,12,2)] # Even months
```

Subsetting:
```{r}
tempHiC[tempHiC > 20] 
```

### Mathematical Operations in Vectors

```{r}
min(tempHiC)
max(tempHiC)
mean(tempHiC)
sd(tempHiC)
```

Subsetting (again):

```{r}
tempHiC[tempHiC > mean(tempHiC)] 
```

## Recycling Rule

Suppose a simple model for global warming:

```{r}
hotter <- tempHiC + 10
hotter
hotter-tempHiC
```

It's not going to be linear... let's try a more sensible one:

```{r}
hotter <- tempHiC + c(3,3,2,2,2,1,1,1,1,2,2,3)
hotter
hotter-tempHiC
```

So we can add a scalar and a vector to another vector. What if they're not the same size?

```{r}
hotter <- tempHiC + c(3,2,1)
hotter
hotter-tempHiC
```


The *Recycling Rule* states that the shorter vector values' are recycled. We get warnings when the lenght of the longer is not a multiple of the lenght of the shorter.

```{r}
hotter <- tempHiC + c(3,2,1,0,-1)
hotter
hotter-tempHiC
```

## Missing Data in Vectors

```{r}
avgTempSJC <- c(22.2,22.4,21.6,19.6,17,16.1,NA,17.1,18.8,19.4,20.3,21.4)
names(avgTempSJC) <- c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")
avgTempSJC
```

Problems with some operators when some values are NAs:
```{r}
min(avgTempSJC)
max(avgTempSJC)
mean(avgTempSJC)
sd(avgTempSJC)
```

We need to be clear about the existence of NAs:
```{r}
min(avgTempSJC,na.rm = TRUE)
max(avgTempSJC,na.rm = TRUE)
mean(avgTempSJC,na.rm = TRUE)
sd(avgTempSJC,na.rm = TRUE)
```

### Filtering by NAs
```{r}
is.na(avgTempSJC)
!is.na(avgTempSJC)
avgTempSJC[is.na(avgTempSJC)]
avgTempSJC[!is.na(avgTempSJC)]
```

