---
title: "Data Frames"
output:   
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data Frames are Tables

Data Frames are the most common way to represent data in R in Data Science and Data Mining applications.

We can create data frames from sets of observations. This below is a column-based example.

```{r}
maxTempSJC <- c(29.7,30.1,29.5,27.3,25.1,24.3,24.1,26.2,27.2,27.3,28,28.7)
avgTempSJC <- c(22.2,22.4,21.6,19.6,17,16.1,15.6,17.1,18.8,19.4,20.3,21.4)
minTempSJC <- c(16.2,16.5,15.7,13.2,10.1,8.9,8.2,9.9,11.9,13.4,14.2,15.3)
months <- c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")
sjcTemps <- data.frame(Max=maxTempSJC,Avg=avgTempSJC,Min=minTempSJC,row.names = months)
sjcTemps
class(sjcTemps)
```
Now for a row-based example, starting with an empty Data Frame:

```{r}
sjcTemps2 <- data.frame(Max=numeric(),Avg=numeric(),Min=numeric())
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=29.7,Avg=22.2,Min=16.2))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=30.1,Avg=22.4,Min=16.5))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=29.5,Avg=21.6,Min=15.7))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=27.3,Avg=19.6,Min=13.2))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=25.1,Avg=17,  Min=10.1))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=24.3,Avg=16.1,Min= 8.9))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=24.1,Avg=15.6,Min= 8.2))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=26.2,Avg=17.1,Min= 9.9))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=27.2,Avg=18.8,Min=11.9))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=27.3,Avg=19.4,Min=13.4))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=28,  Avg=20.3,Min=14.2))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=28.7,Avg=21.4,Min=15.3))
row.names(sjcTemps2) <- months
sjcTemps2
class(sjcTemps2)
```

### Information on Data Frames

Basic functions to get information on data frames:
```{r}
nrow(sjcTemps)
ncol(sjcTemps)
head(sjcTemps)
```

More information with <tt>str</tt>:
```{r}
str(sjcTemps)
```
Even more information (statistics) with <tt>summary</tt>:
```{r}
summary(sjcTemps)
```

### Accessing Data Frames Elements

Accessing by columns' names

```{r}
sjcTemps$Min
sjcTemps$Max
```

Accessing individual values

```{r}
sjcTemps$Min[2]
sjcTemps$Max[2]
```

Basic slicing:

```{r}
sjcTemps[,2]
sjcTemps[2,]
sjcTemps[2,2]
```

Using names:

```{r}
sjcTemps["Jul",]
sjcTemps["Jul",1]
sjcTemps["Jul","Min"]
```

Using ranges, etc.:

```{r}
sjcTemps[1:6,c(1,3)]
sjcTemps["Jul",2:3]
sjcTemps[c("Jan","Jul"),c("Max","Min")]
```

## Reading Data Frames

Creating Data Frames in code is not practical. Let's read them from files.

### From CSV


## Writing and Reading R Data

dump
source



