---
title: "Data Frames"
output:   
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data Frames are Tables

Data Frames are the most common way to represent data in R in Data Science and Data Mining applications.

We can create data frames from sets of observations. This below is a column-based example.

```{r}
maxTempSJC <- c(29.7,30.1,29.5,27.3,25.1,24.3,24.1,26.2,27.2,27.3,28,28.7)
avgTempSJC <- c(22.2,22.4,21.6,19.6,17,16.1,15.6,17.1,18.8,19.4,20.3,21.4)
minTempSJC <- c(16.2,16.5,15.7,13.2,10.1,8.9,8.2,9.9,11.9,13.4,14.2,15.3)
months <- c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")
sjcTemps <- data.frame(Max=maxTempSJC,Avg=avgTempSJC,Min=minTempSJC,row.names = months)
sjcTemps
class(sjcTemps)
```
Now for a row-based example, starting with an empty Data Frame:

```{r}
sjcTemps2 <- data.frame(Max=numeric(),Avg=numeric(),Min=numeric())
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=29.7,Avg=22.2,Min=16.2))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=30.1,Avg=22.4,Min=16.5))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=29.5,Avg=21.6,Min=15.7))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=27.3,Avg=19.6,Min=13.2))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=25.1,Avg=17,  Min=10.1))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=24.3,Avg=16.1,Min= 8.9))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=24.1,Avg=15.6,Min= 8.2))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=26.2,Avg=17.1,Min= 9.9))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=27.2,Avg=18.8,Min=11.9))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=27.3,Avg=19.4,Min=13.4))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=28,  Avg=20.3,Min=14.2))
sjcTemps2 <- rbind(sjcTemps2,data.frame(Max=28.7,Avg=21.4,Min=15.3))
row.names(sjcTemps2) <- months
sjcTemps2
class(sjcTemps2)
```

### Information on Data Frames

Basic functions to get information on data frames:
```{r}
nrow(sjcTemps)
ncol(sjcTemps)
head(sjcTemps)
```

More information with <tt>str</tt>:
```{r}
str(sjcTemps)
```
Even more information (statistics) with <tt>summary</tt>:
```{r}
summary(sjcTemps)
```

### Accessing Data Frames Elements

Accessing by columns' names

```{r}
sjcTemps$Min
sjcTemps$Max
```

Accessing individual values

```{r}
sjcTemps$Min[2]
sjcTemps$Max[2]
```

Basic slicing:

```{r}
sjcTemps[,2]
sjcTemps[2,]
sjcTemps[2,2]
```

Using names:

```{r}
sjcTemps["Jul",]
sjcTemps["Jul",1]
sjcTemps["Jul","Min"]
```

Using ranges, etc.:

```{r}
sjcTemps[1:6,c(1,3)]
sjcTemps["Jul",2:3]
sjcTemps[c("Jan","Jul"),c("Max","Min")]
```

If we can read these values, we can change them.

```{r}
sjcTemps3 <- sjcTemps
sjcTemps3["Jul",] <- 0
sjcTemps3[,"Avg"] <- 0
sjcTemps3["Aug",] <- c(1,2,3)
sjcTemps3["Jul",] <- sjcTemps3["Jul",]+c(3,4)
sjcTemps3["Dec","Max"] <- 50
sjcTemps3
```

### Adding More Data, Removing Data

We can add new columns to a Data Frame: just choose a new column name and an operation to fill it.
```{r}
sjcTemps$Range <- sjcTemps$Max-sjcTemps$Min
sjcTemps$Flag <- TRUE
sjcTemps$Vacations <- c(TRUE,TRUE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE,FALSE,FALSE,FALSE,TRUE)
sjcTemps$Season <- factor(c('Summer','Summer','Autumn','Autumn','Autumn','Winter',
                     'Winter','Winter','Spring','Spring','Spring','Summer'))
sjcTemps
str(sjcTemps)
```

We can also drop columns from a Data Frame:
```{r}
sjcTempsModified <- sjcTemps # avoid changing this one...
sjcTempsModified$Avg <- NULL
sjcTempsModified$Flag <- NULL
head(sjcTempsModified)
```

Other way do to this:
```{r}
sjcTempsModified <- subset(sjcTemps, select = c(Min,Max,Season))
# Also: sjcTempsModified <- sjcTemps[keep]
head(sjcTempsModified)
```

Another one:
```{r}
sjcTempsModified <- subset(sjcTemps, select = -c(Range,Flag,Avg,Season))
head(sjcTempsModified)
```

## Reading and Writing Data Frames

Creating Data Frames in code is not practical. Let's read them from files.

### .. as CSV

Let's read a file like this into R:
<pre>
TAUBATE,Maximum Absolute Temperature,Maximum Average Temperature,Average Temperature,Minimum Average Temperature,Minimum Absolute Temperature,Rainfall (mm),Days with Rainfall >= 1mm,Relative Humidity,Sunshine Hours
Jan,36.7,30.2,23.3,17.7,10.9,233.5,17,76.8,170
Feb,40.1,30.5,23.5,17.8,12.6,192.1,14,76.7,168.2
Mar,36.2,29.9,22.9,17.8,11.6,173.5,12,76.8,179.5
Apr,34.2,27.8,20.8,16.1,5.4,67.1,7,75.9,177.3
May,32.6,25.4,18.1,14.2,4.2,40.9,4,76.2,176.1
Jun,31,24.6,16.7,12.9,1.1,29.3,4,75.7,175.6
Jul,32.1,24.4,16.4,12.6,0.9,31.1,4,72.5,188.9
Aug,35.6,26.4,18.2,14,2.6,41.7,4,69.1,186.2
Sep,37.8,27.6,20,15.7,3.8,64,6,69.2,158.8
Oct,38,27.5,20.7,16.4,7.8,132.8,11,74,148.8
Nov,37.4,28.7,21.7,16.9,8.6,146.2,12,73.8,161.5
Dec,37.5,29.1,22.7,17.1,11,244.6,16,76.8,155
Year,40.1,27.7,20.4,15.8,0.9,1396.8,111,74.5,2045.9
</pre>


```{r}
taubateWeather <- read.csv(file="../Data/Taubate.csv", header=TRUE, sep=",")
taubateWeather
```

Check the formatting of the CSV file: the first column should be named "Month" (see also its last row...)

```{r}
colnames(taubateWeather)[1] <- "Month"
```

Everything seems to be there! Just for fun let's see some statistics (dropping the measure for Year first)
```{r}
taubateWeather <- taubateWeather[taubateWeather$Month != "Year", ]
str(taubateWeather)
```

Let's save it.

```{r}
write.csv(taubateWeather, file = "../Data/Taubate-Fixed.csv")
```
### Writing and Reading R Data

dump
source



