---
title: "Date and Time"
output:   
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Date and Time 

### Creating 

"Now":

```{r}
today <- Sys.Date()
today
class(today)
now <- Sys.time()
now
class(now)
```

### Converting from strings

Dates in format YMD are easy:

```{r}
monday <- as.Date("2017-07-03")
monday
class(monday)
```

Other formats can be parsed.
```{r}
aDate <- "04/07/2017"
tuesday <- as.Date(aDate, "%d/%m/%Y")
tuesday
class(tuesday)
```


<tt>as.Date</tt> ignores time (or does it?)

```{r}
mondayMorning <- as.Date("2017-07-03 05:30:00")
mondayMorning
class(mondayMorning)
```

Converting date and time with <tt>strptime</tt>:

```{r}
aDate <- "04/07/2017"
tuesday <- strptime(aDate, "%d/%m/%Y")
tuesday
class(tuesday)
tuesdayMorning <- strptime("04/07/2017 at 05:20:17", "%d/%m/%Y at %H:%M:%S")
tuesdayMorning
class(tuesdayMorning)
```
### Extracting fields from date and time

```{r}
tuesdayMorning
tM <- as.POSIXlt(tuesdayMorning)
tM$sec
tM$wday
```
You can get fields for
sec, 
min,
hour,
mday (day of the month),
mon (months after the first of the year),
year (since 1900),
wday (0-6, 0=Sunday),
yday (day of year),
isdst (DST).

## Operations with date and time

```{r}
birthEAP <- as.Date("1809/01/19")
today <- Sys.Date()
diff <- today-birthEAP
diff
class(diff)
```

```{r}
today <- Sys.Date()
future <- today+100 # days
future
class(future)
```


There are more flexible ways: StockOverflow is your friend!
https://stackoverflow.com/questions/14454476/get-the-difference-between-dates-in-terms-of-weeks-months-quarters-and-years